Classification structurée pour l'apprentissage par renforcement

### 2012-09-14 klein2012structured_RIA/scirl_ria.tex
Traduction de la première section

\subsection{Classification with Linearly Parameterized Score
Functions} \label{subsec:background:classif}

Let $\X$ be a compact or a finite set (of inputs to be classified)
and let $\Y$ be a finite set (of labels). Assume that inputs $x\in
\X$  are drawn according to some unknown distribution $\prob(x)$ and
that there exists some oracle which associates to each of these
inputs a label $y\in Y$ drawn according to the unknown conditional
distribution $\prob(y|x)$. Generally speaking, the goal of
multi-class classification is, given a training set
$\{(x_i,y_i)_{1\leq i \leq N}\}$ drawn according to $\prob(x,y)$, to
produce a decision rule $g\in\Y^\X$ which aims at minimizing the
classification error $\E[\chi_{\{g(x)\neq y\}}] = \prob(g(x)\neq
y)$, where $\chi$ denotes the indicator function.

Here, we consider a more restrictive set of classification
algorithms. We assume that the decision rule associates to an input
the argument which maximizes a related score function, this score
function being linearly parameterized et the associated parameters
being learnt by the algorithm. More formally, let $\psi(s,a) =
(\psi_1(x,y)  \dots  \psi_d(x,y))^\top\in \mathbb{R}^d$
%\begin{pmatrix}   \psi_1(x,y) & \dots & \psi_d(x,y)
%\end{pmatrix}^\top \in \mathbb{R}^d$
be a feature vector whose components are $d$ basis functions
$\psi_i\in\mathbb{R}^{\X\times\Y}$. The linearly parameterized score
function $s_w\in\mathbb{R}^{\X\times \Y}$ of parameter vector
$w\in\mathbb{R}^d$ is defined as $s_w(x,y) = w^\top \psi(x,y)$. The
associated decision rule $g_w\in{\Y^\X}$ is defined as $g_w(x) \in
\argmax_{y\in\Y}s_w(x,y)$. Using a training set $\{(x_i,y_i)_{1\leq
i\leq N}\}$, a linearly parameterized score function-based
multi-class classification (MC$^2$ for short) algorithm computes a
parameter vector $\theta_c$. The quality of the solution is
quantified by the classification error $\epsilon_c =
\prob(g_{\theta_c}(x)\neq y)$.

We do not consider a specific MC$^2$ algorithm, as long as it
classifies inputs by maximizing the argument of a linearly
parameterized score function. For example, one may choose a
multi-class support vector machine~\cite{Guermeur:2007} (taking the
kernel induced by the feature vector) or a structured large margin
approach~\cite{Taskar:2005}. Other choices may be possible, one can
choose its preferred algorithm.

\section{Structured Classification for Inverse Reinforcement
Learning} \label{sec:scirl}

\subsection{General Algorithm}
\label{subsec:scirl:algo}

Consider the classification framework of
Sec.~\ref{subsec:background:classif}. The input $x$ may be seen as a
state et the label $y$ as an action. Then, the decision rule
$g_w(x)$ can be interpreted as a policy which is greedy according to
the score function $w^\top \psi(x,y)$, which may itself be seen as
an action-value function. Making the parallel with
Eq.~\eqref{eq:def:mu}, if $\psi(x,y)$ is the feature expectation of
some policy $\pi$ which produces labels of the training set, et if
the classification error is small, then $w$ will be the parameter
vector of a reward function for which we may hope the policy $\pi$
to be near optimal. Based on these remarks, we're ready to present
the proposed Structured Classification-based IRL (SCIRL) algorithm.

>>>>

\subsection{Classifieurs à fonction de score paramétrée linéairement} \label{subsec:background:classif}

Soit $\X$ un ensemble fini ou compact (d'entrées à classifier)
 et soit $\Y$ un ensemble fini de (de labels). Supposons que les entrées $x\in
\X$ sont tirées selon une distribution inconnue $\prob(x)$ et qu'il existe un oracle qui à chacune de ces entrées associe un label $y\in Y$ tiré selon une distribution de probabilité conditionnelle inconnue $\prob(y|x)$. De manière générale, la classification multi-classe cherche étant donné un set d'entraînement $\{(x_i,y_i)_{1\leq i \leq N}\}$ tiré selon $\prob(x,y)$, une règle de décision $g\in\Y^\X$ minimisant l'erreur de classification $\E[\chi_{\{g(x)\neq y\}}] = \prob(g(x)\neq y)$, où $\chi$ est la fonction indicatrice.

Nous nous préoccupons ici d'un ensemble plus réduit de classifieurs. Nous supposons que la règle de décision associe à l'entrée l'argument qui maximise une certaine fonction de score, celle-ci est paramétrée linéairement, les paramètres étant appris par le classifieur. Formellement, soit $\psi(s,a) =
(\psi_1(x,y)  \dots  \psi_d(x,y))^\top\in \mathbb{R}^d$
%\begin{pmatrix}   \psi_1(x,y) & \dots & \psi_d(x,y)
%\end{pmatrix}^\top \in \mathbb{R}^d$
un vecteur d'attributs dont les composantes sont $d$ fonctions de base $\psi_i\in\mathbb{R}^{\X\times\Y}$. La fonction de score linéairement paramétrée $s_w\in\mathbb{R}^{\X\times \Y}$ de vecteur de paramètres $w\in\mathbb{R}^d$ est définie par $s_w(x,y) = w^\top \psi(x,y)$. La règle de décision associée $g_w\in{\Y^\X}$ est définie par $g_w(x) \in \argmax_{y\in\Y}s_w(x,y)$. En se basant sur un set d'entraînement $\{(x_i,y_i)_{1\leq
i\leq N}\}$, un classifieur multi-classe (noté MC$^2$ pour \emph{Multi-Class Classifier}) à fonction de score paramétrée linéairement choisit un vecteur de paramètres $\theta_c$. La qualité de ce choix est quantifiée par l'erreur de classification $\epsilon_c =
\prob(g_{\theta_c}(x)\neq y)$.

Notre étude est valable pour tout algorithme de classification tant qu'il opère en maximisant l'argument d'une fonction de score paramétrée linéairement. Par exemple, une machine à vecteur support multi-classe pourrait être choisie~\cite{Guermeur:2007} (en prenant le noyau induit par le vecteur de paramètres) de même qu'une approche à marge vaste et structurée~\cite{Taskar:2005}. D'autres choix sont possibles, chacun peut choisir son algorithme fétiche.

\section{Classification structurée pour l'apprentissage par renforcement inverse} \label{sec:scirl}

\subsection{Forme générale de l'algorithme}
\label{subsec:scirl:algo}

On se place dans le cadre de la classification décrit en Sec.~\ref{subsec:background:classif}. L'entrée $x$ oeut être vue comme un état et le label $y$ comme une action. Il s'ensuit que l'on peut interpréter la règle de décision 
$g_w(x)$ comme une politique gloutonne vis-à-vis de la fonction de score $w^\top \psi(x,y)$, qui peut elle même être vue comme une fonction de qualité. Making the parallel with
Eq.~\eqref{eq:def:mu}, if $\psi(x,y)$ is the feature expectation of
some policy $\pi$ which produces labels of the training set, et if
the classification error is small, then $w$ will be the parameter
vector of a reward function for which we may hope the policy $\pi$
to be near optimal. Based on these remarks, we're ready to present
the proposed Structured Classification-based IRL (SCIRL) algorithm.
